The UK Biobank Study: Systematic Review of the Reporting Quality in UK Biobank

### 2021-11-05 manuscript/manuscript.Rmd
replace table1 with table2, don't recode unsure because should all be resolved, reorder bar labels, format kappa tables and add all NA cols, create supp table 2, add narrative results

############
# libraries ####
#############
library(magrittr)
library(ggplot2)
library(tableone)
library(gtools)
library(tidyr)
library(RColorBrewer)
library(dplyr)
library(irr)
library(lpSolve)

@@

strobe_stem <- gsub("_.*", "", strobe_cols_names) %>%
  unique()

###################################
# split into strobe and non-strobe ####
#################################

@@

# save version of not strobe df without qualtrics metadata variables to create table 1

create_table1_df <- function(){
  # save pattern to find qualtrics variables
  qual <- c("distribution_channel", "duration_in_seconds", "end_date", "finished", "progress","recorded_date", "response_id", "response_type", "start_date", "user_language")
  qual_pat <- paste0("^", qual, ".", collapse = "|")

  table1 <- not_s_df[, -qual_cols]

  # drop article_id col as don't need freqs of these
  table1 <- table1[, -which(colnames(table1) %in% "article_id")]

  #convert all non-strobe cols into characters as easier than setting multiple different factor levels
  for(i in colnames(table1)){
    table1[[i]] <- as.character(table1[[i]])
  }

  # ensure UK and USA country values are capitalised as frequency is case-sensitive
  uk_usa <- grep("^uk$|^usa$", table1$country, ignore.case = T)
  table1$country[uk_usa] <- toupper(table1$country[uk_usa])
  return(table1)
}

table1 <- create_table1_df()

CreateTableOne(data = table1, includeNA = F)%>%
  print(., noSpaces = T, showAllLevels = T) %>%
  # export
  write.csv(., "../outputs/table1_all_levels.csv")

CreateTableOne(data = table1, includeNA =F )%>%
  print(., noSpaces = T) %>%
  # export
  write.csv(., "../outputs/table1_no_levels.csv")

@@

# recode strobe variables into factors so can analyse later
factorise_strobe <- function(){
  # check strobe cols correct
  if(!all(strobe_cols_names %in% colnames(s_df))) stop("some strobe cols not in s_df")
  # save s_df before recoding factors
  pre_recode <- s_df
  # recode all strobe cols in s_df
  for(i in strobe_cols_names){
    # set all cols to factor
    s_df[[i]] <- as.factor(s_df[[i]])
    #recode all factor levels
    levels(s_df[[i]]) <- list(Yes = "Yes", PartiallyExternal = "PartiallyExternal", Partially = "Partially", No = "No", Unsure = "Unsure", na = "na")
  }

@@

  # check all are characters
  for(i in id){
    if(is.character(x[, i]) == F) stop(i, "isn't a character")
  }
  # recode into binary
  if(style == "binary"){
    for(i in id){
      x[, i] <-  gsub("^Yes$", "1", x[, i]) %>%
        gsub("^No$|^Partially$|^Unsure$|^PartiallyExternal$", "0", .) %>%
        gsub("na", NA, .) %>%
        as.numeric(.)
      if(is.numeric(x[, i]) == F) stop(i, "not numeric")
    }
  }
  # recode into ternary
  if(style == "ternary") {
    for(i in id){
      x[, i] <-  gsub("^Yes$|^PartiallyExternal$", "1", x[, i]) %>%
        gsub("^Partially$", "0.5", .) %>%
        gsub("^No$|^Unsure$", "0", .) %>%
        gsub("na", NA, .) %>%
        as.numeric(.)
      if(is.numeric(x[, i]) == F) stop(i, "not numeric")
    }
  }
  #check strobe cols
  if(!all(strobe_cols_names %in% colnames(x))) stop("some strobe cols not in s_df_tri")

@@

  bar_data <- tidyr::pivot_longer(bar_data, colnames(bar_data), names_to = "strobe_item", values_to = "value") 
  x <- table(bar_data$strobe_item, bar_data$value, useNA = "always")
  y <- prop.table(table(bar_data$strobe_item, bar_data$value), 1) %>%
    round(., 2)
  # convert to dataframe
  x <- as.data.frame.matrix(x)
  y <- as.data.frame.matrix(y)

@@

bar_data$response <- as.factor(bar_data$response)
levels(bar_data$response) <- c("No", "Partially", "Yes", "Not Applicable")

@@

# strobe item on x axis and plot % yes
bar_data %>%ggplot(aes(x=strobe_item, y=count, fill = response)) + 
  # blue fill
  geom_bar(stat="identity") +
  theme_classic() +
  theme(text = element_text(size=30))+
  scale_x_discrete(labels= labels$x_labels) +
  scale_y_continuous(breaks = seq(0, nrow(df), by = 10)) + 
  coord_flip() +
  xlab("STROBE Item") +
  ylab("Count") +
  ggtitle("Figure 1. STROBE completion")+
  scale_fill_brewer(palette = "Spectral") +
  #geom_text(aes(label = count), vjust = 0, size = 8) +
  theme(plot.title = element_text(size = 22)) 
dev.off()

@@

  kappa <- c()
  z <- c()
  p_value <- c()
  variable <- c()

  for(i in stem){
    x <- df[, grep(i, colnames(df))]
    y <- kappa2(x, weight = "unweighted")
    variable <- c(variable, i)
    kappa <- c(kappa, y$value)
    z <- c(z, y$statistic)
    p_value <- c(p_value, y$p.value)
  }

  irr_df <- data.frame(variable = variable, kappa= kappa, z= z, p_value = p_value)

}

@@

inter_na <- divide_na()
na_kappa <- create_kappa_df(inter_na)

@@

@@

Exclusion criteria:
* Article findings were produced from analysis(es) of pooled data from multiple studies.
These are studies which have used data from participants from the UKB and other datasets in the same analysis.
* Meta-analyses; narrative and systematic reviews; preprints; post-prints; replies; letters to the editor; corrections; and any other publication types that are not full research articles.
* Retracted articles

@@

## Reporting quality

![Figure 1](Rplot9.png)

# Discussion

@@

>>>>

############
# libraries ####
#############
library(magrittr)
library(ggplot2)
library(tableone)
library(gtools)
library(tidyr)
library(RColorBrewer)
library(dplyr)
library(irr)
library(lpSolve)
library(kableExtra)
library(pander)

@@

strobe_stem <- gsub("_.*", "", strobe_cols_names) %>%
  unique()
###################################
# split into strobe and non-strobe ####
#################################

@@

@@

# recode strobe variables into factors so can analyse later
factorise_strobe <- function(){
  # check strobe cols correct
  if(!all(strobe_cols_names %in% colnames(s_df))) stop("some strobe cols not in s_df")
  # save s_df before recoding factors
  pre_recode <- s_df
  # recode all strobe cols in s_df
  for(i in strobe_cols_names){
    # set all cols to factor
    s_df[[i]] <- as.factor(s_df[[i]])
    #recode all factor levels
    levels(s_df[[i]]) <- list(Yes = "Yes", PartiallyExternal = "PartiallyExternal", Partially = "Partially", No = "No", na = "na")
  }

@@

  # check all are characters
  for(i in id){
    if(is.character(x[, i]) == F) stop(i, "isn't a character")
  }
  # recode into binary
  if(style == "binary"){
    for(i in id){
      x[, i] <-  gsub("^Yes$", "1", x[, i]) %>%
        gsub("^No$|^Partially$|^PartiallyExternal$", "0", .) %>%
        gsub("na", NA, .) %>%
        as.numeric(.)
      if(is.numeric(x[, i]) == F) stop(i, "not numeric")
    }
  }
  # recode into ternary
  if(style == "ternary") {
    for(i in id){
      x[, i] <-  gsub("^Yes$|^PartiallyExternal$", "1", x[, i]) %>%
        gsub("^Partially$", "0.5", .) %>%
        gsub("^No$", "0", .) %>%
        gsub("na", NA, .) %>%
        as.numeric(.)
      if(is.numeric(x[, i]) == F) stop(i, "not numeric")
    }
  }
  #check strobe cols
  if(!all(strobe_cols_names %in% colnames(x))) stop("some strobe cols not in s_df_tri")

@@

  bar_data <- tidyr::pivot_longer(bar_data, colnames(bar_data), names_to = "strobe_item", values_to = "value") 
  x <- table(bar_data$strobe_item, bar_data$value, useNA = "always") %>%
    # convert to dataframe
    as.data.frame.matrix()
  y <- prop.table(table(bar_data$strobe_item, bar_data$value), 1) %>%
    # round so more interpretable
    round(., 2) %>%
    as.data.frame.matrix()

@@

bar_data$response <- as.factor(bar_data$response)

# set levels using list so is accurate
levels(bar_data$response) <- list(NotApplicable = "na", No = 0, Partially = 0.5, Yes = 1)

# add space to NA level since can't do this in list
levels(bar_data$response)[levels(bar_data$response) == "NotApplicable"] <- "Not Applicable"

@@

# strobe item on x axis and plot % yes
bar_data %>%ggplot(aes(x=strobe_item, y=count, fill = response)) + 
  # blue fill
  geom_bar(stat="identity") +
  theme_classic() +
  theme(text = element_text(size=30))+
  scale_x_discrete(labels= labels$x_labels) +
  scale_y_continuous(breaks = seq(0, nrow(df), by = 10)) + 
  coord_flip() +
  xlab("STROBE Item") +
  ylab("Count") +
  ggtitle("Figure 1. STROBE completion")+
  scale_fill_brewer(palette = "Greys") +
  #geom_text(aes(label = count), vjust = 0, size = 8) +
  theme(plot.title = element_text(size = 22)) 
dev.off()

@@

  # save stems of duplicate and not applicable cols so can join back in
  na_items <- c("5_vi", "6a_iii", "10", "8starred_iii")
  na_stem <-  grep(paste0(na_items, collapse ="|"), stem, value =T)

  # remove all stems that are duplicate or not applicable as will throw error in kappa2 if set "NA" to missing and all are set to NA after qualtrics

  stem <- stem[-which(stem == na_stem)]
  #save empty objects so can save outputs from for loop
  kappa <- c()
  z <- c()
  p_value <- c()
  variable <- c()
  n_match <- c()
  n_conflicts <-c()
  n_na <- c()

  for(i in stem){
    x <- df[, grep(i, colnames(df))]
    y <- kappa2(x, weight = "unweighted")
    variable <- c(variable, i)
    kappa <- c(kappa, y$value)
    z <- c(z, y$statistic)
    p_value <- c(p_value, y$p.value)
    n_match <- c(n_match, sum(x[[1]] == x[[2]], na.rm = T))
    n_conflicts <- c(n_conflicts, sum(x[[1]] != x[[2]], na.rm = T))
    # count number of items with NA--NA dyads or notNA--NA dyads so are excluded from kappa
    n_na <- c(n_na, sum(is.na(x[[1]] != x[[2]])))
  }

  # bind objects into df so can render in manuscript
  irr_df <- data.frame(STROBEItem = variable, Kappa = kappa, Zstatistic = z, pvalue = p_value, Matches = n_match, Conflicts = n_conflicts, NotApplicableUnsure = n_na)

  # create df of empty varibles and bind to main df so will appear in table
  empty <- data.frame(STROBEItem = na_stem)
  irr_df <- full_join(irr_df, empty, by = "STROBEItem")

  # round all numeric cols so more interpretable
  nums <- colnames(irr_df)[colnames(irr_df) != "STROBEItem"]
  for(i in nums){
    x <- irr_df[[i]]
    if(!is.numeric(x)) stop(i, " col not numeric")
    x <- round(x, digits = 2)
    irr_df[[i]] <- x
  }

  # remove backslash that was added to variable names
  irr_df$STROBEItem <- gsub("\\\\.*", "", irr_df$STROBEItem)

  # reorder
  irr_df <- irr_df[gtools::mixedorder(irr_df$STROBEItem), ]

  # set rownames to null to ensure none created
  row.names(irr_df) <- NULL

  return(irr_df)
}

@@

inter_na <- divide_na()
kappa_na <- create_kappa_df(inter_na)

@@

kappa_item$STROBEItem <- gsub("^s|_.*|starred", "", kappa_item$STROBEItem)

@@

Exclusion criteria:

* Article findings were produced from analysis(es) of pooled data from multiple studies.
These are studies which have used data from participants from the UKB and other datasets in the same analysis.
* Meta-analyses; narrative and systematic reviews; preprints; post-prints; replies; letters to the editor; corrections; and any other publication types that are not full research articles.
* Retracted articles

@@

Of the `r nrow(df)` included studies `r sum(grepl("Case-control", df$designs))` contained case-control designs, `r sum(grepl("Cohort", df$designs))` cohort, and `r sum(grepl("Cross-sectional", df$designs))` cross-sectional. These numbers add up to more than 70 because some articles contained cross-sectional and cohort studies.
UK based researchers led the vast majority with `r sum(grepl("UK", df$country, ignore.case = F))` having authors affiliated with institutions in the UK.
`r sum(grepl("2014", df$year))` studies were published in 2014, `r sum(grepl("2015", df$year))` in 2015, `r sum(grepl("2016", df$year))` in 2016, `r sum(grepl("2017", df$year))` in 2017, `r sum(grepl("2018", df$year))` in 2018, and `r sum(grepl("2019", df$year))` in 2019.

## STROBE Reporting quality

Using the data extraction form on Qualtrics, we assessed reporting quality separately for all subdivisions of the STROBE items and for the “starred” items associated with any STROBE items that had an asterisk (*) next to it. 
For the final analysis, we excluded all starred subdivisions because inter-rater reliability was so poor and the applicability of the items was often uncertain. 
We then summed the subdivisions to indicate the reporting quality of each STROBE item, these results are presented in Figure 1. 
For completeness, the results for the subdivisions are available in Supplementary Table 2.

We originally assessed reporting quality separately for all subdivisions of the STROBE items.
The results for the subdivisions are reported in Supplementary Table 2.

![](Rplot9.png)

We calculated kappa statistics to estimate the inter-rater reliability for each of the STROBE items, these are presented in Table 1.

```{r table1}

doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')

if(doc.type == "docx"){pander(kappa_item, caption = "Table 1. Inter-rater reliability for the STROBE items")} else {kable(kappa_item)}

## Other reporting quality

We collected reporting quality data in addition to the STROBE items.
We intended to conduct exploratory analyses using these variables, for example comparing articles that reported using some form of reporting guidelines to those who did not.
We did not conduct an exploratory analyses into the relationships between these variabels because subgroup analyses for all of these variables would have very small subgroup samples. 
Instead we report the descriptive statistics for each variable in Table 2 below.

*Table 2. Descriptive statistics for additional reporting quality variables*

```{r table2, echo=F, include = F}

# save version of not strobe df without qualtrics metadata variables to create table 2

create_table2 <- function(){
  # save pattern to find qualtrics variables
  qual <- c("distribution_channel", "duration_in_seconds", "end_date", "finished", "progress","recorded_date", "response_id", "response_type", "start_date", "user_language")
  qual_pat <- paste0("^", qual, ".", collapse = "|")

  table2 <- not_s_df[, -qual_cols]

  # drop cols that are qualitative so contain unique info for each observation so no point in calculating frequencies
  table2 <- table2[, -which(colnames(table2) %in% c("article_id", "correction_address", "open_access_address", "reg_id", "correction_address"))]

  #convert all non-strobe cols into characters as easier than setting multiple different factor levels
  for(i in colnames(table2)){
    table2[[i]] <- as.character(table2[[i]])
  }

  # ensure UK and USA country values are capitalised as frequency is case-sensitive
  uk_usa <- grep("^uk$|^usa$", table2$country, ignore.case = T)
  table2$country[uk_usa] <- toupper(table2$country[uk_usa])

  # drop cols that are characteristics
  chars <- c("designs", "country", "year")
  if(!all(chars %in% colnames(table2))) stop("missing char elements in colnames")
  table2 <- table2[, -which(colnames(table2) %in% chars)]

  # set missing & na as won't make a difference to absolute numbers
  table2[is.na(table2)|table2 == "na"] <- "Not Applicable"
  table2[table2 == "ni"] <- "No information"
  table2[table2 == "TRUE"] <- "Yes"
  table2[table2 == "FALSE"] <- "No"

  # rename table2 colnames so will come out as full words in createtableon
  colnames(table2)[colnames(table2) == "code"] <- "Report shared analysis code"
  colnames(table2)[colnames(table2) == "exposure"] <- "Exposures reported as exposures"
  colnames(table2)[colnames(table2) == "outcome"] <- "Outcomes reported as outcomes"
  colnames(table2)[colnames(table2) == "reg"] <- "Report analyses registered"
  colnames(table2)[colnames(table2) == "strobe"] <- "Report using STROBE guidelines"
  colnames(table2)[colnames(table2) == "strobe_cite"] <- "Cite STROBE guidelines if used"
  colnames(table2)[colnames(table2) == "ukb_credit"] <- "Report UK Biobank credit statement correctly"
  colnames(table2)[colnames(table2) == "var_id"] <- "Report UK Biobank variable IDs"
  colnames(table2)[colnames(table2) == "other_guidelines"] <- "Report using reporting guidelines other than STROBE"
  colnames(table2)[colnames(table2) == "open_access_publisher"] <- "Published as open access"
  colnames(table2)[colnames(table2) == "correction"] <- "Has a correction"
  colnames(table2)[colnames(table2) == "ukb_app_pres"] <- "Report UK Biobank application number" 
  colnames(table2)[colnames(table2) == "email_pres"] <- "Has email address for corresponding author"
  colnames(table2)[colnames(table2) == "keywords_pres"] <- "Has keywords"
  colnames(table2)[colnames(table2) == "coi_pres"] <- "Report conflict of interest statement"
  colnames(table2)[colnames(table2) == "access_supp"] <- "Has supplementary material"

# drop retraction col because none should be
  table2 <- table2[, -which(colnames(table2) %in% "retraction")]
  return(table2)
}

table2 <- create_table2()

CreateTableOne(data = table2, includeNA = F) %>%
  print(., noSpaces = T) %>%
  write.csv(., "../outputs/table2.csv")

```
# Discussion

@@

# Supplementary Material

## Supplementary Table 2

```{r SuppTable2}

create_stab2 <- function(){
  stab2 <- df[, which(colnames(df) %in% c(strobe_cols_names))]
  # convert all to characters so can easily reset values and table will just treat as nominal
  for(i in colnames(stab2)){
    stab2[[i]] <- as.character(stab2[[i]])
  }

  # set missing to na as won't make a difference to absolute numbers
  stab2[is.na(stab2)|stab2 == "na"] <- "Not Applicable"
  stab2[stab2 == "PartiallyExternal"] <- "Partially-External"

  # pivot so can table frequencies
  stab2 <- tidyr::pivot_longer(stab2, colnames(stab2), names_to = "strobe_item", values_to = "value") %>%
    # reorder so follows strobe order
    .[mixedorder(as.character(.$strobe_item)), ]

  # remove s from strobe item
  stab2$strobe_item <- gsub("^s", "", stab2$strobe_item)

  # table
  stab2 <- table(stab2$strobe_item, stab2$value, useNA = "ifany")

  return(stab2)
}

stab2 <- create_stab2()

if(doc.type == "docx"){pander(stab2, caption = "Supplementary Table 2. Descriptive statistics for reporting quality of subdivisions of all STROBE items")} else {kable(stab2)}
```
